%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 2.5 (27/8/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Version 2.x major modifications by:
% Vel (vel@latextemplates.com)
%
% This template is based on a template by:
% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% Template license:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
12pt, % The default document font size, options: 10pt, 11pt, 12pt
%oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
english, % ngerman for German
singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
%liststotoc, % Uncomment to add the list of figures/tables/etc to the table of contents
%toctotoc, % Uncomment to add the main table of contents to the table of contents
%parskip, % Uncomment to add space between paragraphs
%nohyperref, % Uncomment to not load the hyperref package
headsepline, % Uncomment to get a line under the header
%chapterinoneline, % Uncomment to place the chapter title next to the number on one line
consistentlayout, % Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
]{MastersDoctoralThesis} % The class file specifying the document structure


\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters

\usepackage{mathpazo} % Use the Palatino font by default

\usepackage[backend=biber,% style=authoryear,
natbib=true]{biblatex}
\addbibresource{refs.bib} % The filename of the bibliography

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\usepackage{pifont}

\usepackage{cover} % Styles for the cover (title) page

\input{packages}
\input{macros}

%----------------------------------------------------------------------------------------
%	GENERAL SETTINGS
%----------------------------------------------------------------------------------------

\setlength{\hfuzz}{20pt}
\renewcommand{\labelitemi}{--}

%----------------------------------------------------------------------------------------
%	MARGIN SETTINGS
%----------------------------------------------------------------------------------------

\geometry{
	paper=a4paper, % Change to letterpaper for US letter
	inner=1.5cm, % Inner margin
	outer=1.5cm, % Outer margin
	bindingoffset=.5cm, % Binding offset
	top=1.5cm, % Top margin
	bottom=1.5cm, % Bottom margin
	%showframe, % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%----------------------------------------------------------------------------------------

\thesistitle{Vérification formelle d'une méthodologie pour la conception et la production de systèmes numériques critiques} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Dr. James \textsc{Smith}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{Doctor of Philosophy} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Vincent \textsc{Iampietro}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

\subject{Biological Sciences} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\university{\href{http://www.university.com}{University Name}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\department{\href{http://department.university.com}{Department or School Name}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\group{\href{http://researchgroup.university.com}{Research Group Name}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\faculty{\href{http://faculty.university.com}{Faculty Name}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname

\AtBeginDocument{
\hypersetup{pdftitle=\ttitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
}

\begin{document}

% Macros to put after the beginning of the document

\input{docmacros}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title{\protect \parbox{\textwidth}{\centering Vérification formelle
    d'une méthodologie pour la conception et la production de systèmes
    numériques critiques}}

\englishtitle{\protect \parbox{\textwidth}{\centering\it Formal
    verification of a methodology for the design and production of
    safety-critical digital systems}}

\directorA{David Delahaye}
\directorB{David Andreu}
% si deux personnes pour direction these, utiliser directorB et enlever %direction-2% dans fichier couverture.sty
%\directorB{[]}

\date{16 décembre 2021}

\thispagestyle{empty}

\universite{UNIVERSITÉ DE MONTPELLIER}
% \univlogo{Figures/logo_UM.png}
% \univwallpaper{Figures/wallpaper.png}
\doctoral{: Information, Structures, Syst\`emes}
\researchunit{LIRMM}
\specialisation{Informatique}
\jury{Sandrine Blazy}{Professeure}{Université de Rennes}{Rapporteure}
\jury{Frédéric Boniol}{Maître de recherche}{ONERA, Toulouse}{Rapporteur}
\jury{David Déharbe}{Docteur, Ingénieur}{Clearsy}{Examinateur}
\jury{Marc Pouzet}{Professeur}{ENS, Paris}{Examinateur}
\jury{David Delahaye}{Professeur}{Université de Montpellier}{Directeur}
\jury{David Andreu}{MCF}{Université de Montpellier/Neurinnov}{Co-directeur}

\maketitle

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

% \begin{acknowledgements}
%   \addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
%   The acknowledgments and the people to thank go here, don't forget to
%   include your project advisor\ldots
% \end{acknowledgements}

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\setcounter{tocdepth}{2}
\tableofcontents % Prints the main table of contents

% \listoffigures % Prints the list of figures

% \listoftables % Prints the list of tables

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

% \begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)
%   \textbf{SITPN} & \textbf{S}ynchronously executed \textbf{I}nterpreted \textbf{T}ime \textbf{P}etri \textbf{N}et with priorities \\
%   \textbf{VHDL} & \textbf{V}ery high speed integrated circuit \textbf{H}ardware \textbf{D}escription \textbf{L}anguage \\
%   \textbf{CIS} & \textbf{C}omponent \textbf{I}nstantiation \textbf{S}tatement \\
%   \textbf{PCI} & \textbf{P}lace \textbf{C}omponent \textbf{I}nstance \\
%   \textbf{TCI} & \textbf{T}ransition \textbf{C}omponent \textbf{I}nstance \\
%   \textbf{GPL} & \textbf{G}eneric \textbf{P}rogramming \textbf{L}anguage \\
%   \textbf{HDL} & \textbf{H}ardware \textbf{D}escription \textbf{L}anguage \\
%   \textbf{LRM} & \textbf{L}anguage \textbf{R}eference \textbf{M}anual \\
%   \textbf{DSL} & \textbf{D}omain \textbf{S}pecific \textbf{L}anguage \\
%   \textbf{MDE} & \textbf{M}odel-\textbf{D}riven \textbf{E}ngineering \\
%   \textbf{FM} & \textbf{F}ormal \textbf{M}ethods \\
% \end{abbreviations}

%----------------------------------------------------------------------------------------
%	DEDICATION
%----------------------------------------------------------------------------------------

% \dedicatory{For/Dedicated to/To my\ldots} 

%----------------------------------------------------------------------------------------
%	FRENCH ABSTRACT
%----------------------------------------------------------------------------------------

\begin{frenchabstract}
  La production de circuits numériques complexes est devenue
  impossible sans l'aide des ordinateurs. La méthodologie \hilecop{}
  (HIgh LEvel hardware COmponent Programming) assiste les ingénieurs
  dans la conception et la production de circuits numériques dans le
  contexte des systèmes critiques, i.e. systèmes dont le
  malfonctionnement peut résulter en la perte de vies humaines, des
  catastrophes naturelles, des désastres économiques, etc. A titre
  d'exemple, la société
  Neurinnov\footnote{\url{https://neurinnov.com/}} applique la
  méthodologie \hilecop{} pour la production de neuroprothèses,
  considérées comme des dispositifs médicaux hautement critiques par
  la loi de régulation de
  l'UE\footnote{\url{https://eur-lex.europa.eu/eli/reg/2017/745/2020-04-24}}. Dans
  \hilecop{}, les ingénieurs produisent un modèle de circuit
  numérique. Ils utilisent un formalisme graphique qui regroupe les
  diagrammes à composant et un type particulier de réseaux de Petri
  (RdP). Ensuite, le modèle est transformé en une représentation
  textuelle intermédiaire décrite en langage \vhdl{} (Very high speed
  integrated circuit Hardware Description Language). Finalement, un
  compilateur/synthétiseur industriel génère un circuit numérique
  physique, i.e. un ASIC ou sur carte FPGA, depuis la représentation
  \vhdl{}. Ici, l'utilisation des
  RdPs % comme structure sous-jacente aux
  % modèles de circuits numériques
  est liée au contexte des systèmes critiques. Les RdPs permettent la
  vérification de propriétés sur les modèles de circuits numériques
  grâce à l'application de techniques de \textit{model-checking}.
  Cependant, une des transformations décrite dans la méthodologie
  \hilecop{} pourrait altérer le \textit{comportement} (ou
  \textit{sémantique}) des modèles initiaux, invalidant ainsi les
  précédentes étapes de vérification. Le but de cette thèse est de
  prouver que la transformation modèle-vers-texte de \hilecop{}, qui
  génère une description \vhdl{} depuis un modèle de circuit
  numérique, préserve le comportement des modèles d'entrée, i.e.: pour
  tout modèle passé en entrée de la transformation, la description
  \vhdl{} résultante se comporte de la même manière. Pour prouver
  cette propriété, nous nous inspirons des travaux menés sur la
  vérification formelle de compilateurs (notamment sur le compilateur
  C certifié \ccert{} \cite{Leroy2009}). Notre approche est celle de
  la vérification déductive interactive avec assistants de preuve.
  % Spécifiquement, nous voulons prouver cette
  % propriété dans le contexte de la vérification déductive interactive
  % avec assistants de preuve.
  Dans ce contexte, les étapes pour établir la propriété de
  préservation de comportement de la transformation sont: (1)
  formaliser la sémantique d'execution de la représentation source,
  (2) de la représentation cible, (3) décrire la transformation, et
  (4) prouver un théorème de préservation sémantique. Même en suivant
  ce processus clairement détaillé, les spécificités de la
  transformation modèle-vers-texte de \hilecop{} (comparaît notamment
  aux compilateurs) apportent de nouvelles questions recherches et des
  challenges à chaque étape. Dans cette thèse, nous utilisons
  l'assistant à la preuve \coq{} pour nous accompagner tout au long du
  processus. Finalement, nous avons prouvé que la transformation de
  \hilecop{} préserve le comportement de tous modèles initiaux. La
  mécanisation complète de la preuve avec \coq{} est un travail en
  cours.
\end{frenchabstract}

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}
  The complexity of digital hardware circuits makes it difficult to
  produce them without the help of computers.  The \hilecop{} (HIgh
  LEvel hardware COmponent Programming) methodology aims at the
  assistance of engineers in the design and production of such digital
  circuits. The context of production is the one of
  \textit{safety-critical} digital systems, i.e. systems which failure
  could result in direct human losses, natural catastrophes, economic
  disasters, etc. To give an example, the
  Neurinnov\footnote{\url{https://neurinnov.com/}} company leverages
  the \hilecop{} methodology to produce highly critical medical
  devices known as neuroprostheses.  In \hilecop{}, engineers rely on
  a graphical formalism, based on component diagrams and a particular
  kind of Petri nets, to produce a model of a digital circuit. Then, a
  computer program turns the model into an intermediary description
  written in \vhdl{} (Very high speed integrated circuit Hardware
  Description Language). Finally, an industrial compiler/synthesizer
  transforms the \vhdl{} description into a concrete physical circuit
  on an FPGA, or as an ASIC. The use of Petri nets permits the
  engineers to describe a \textit{formal} model of a digital
  circuit. The mathematical foundations of Petri nets enable the use
  of model-checking techniques. Thus, proofs can be brought that the
  produced models verify certain \textit{soundness} properties.
  However, even with a \textit{sound} model of a circuit, one
  transformation step could alter the behavior of the initial
  model. The goal of this thesis is to bring the formal proof that the
  model-to-text transformation from a \hilecop{} high-level model to a
  \vhdl{} description is \textit{semantic preserving} (or
  \textit{behavior preserving}); i.e. for all high-level model given
  as an input to the transformation, the resulting \vhdl{} description
  behaves similarly. To perform this task, we draw our inspiration
  from the works pertaining to the formal verification of compilers
  for programming languages (especially from the certified C compiler
  \ccert{} \cite{Leroy2009}). Specifically, we are interested in
  proving the property of semantic preservation in the context of
  \textit{deductive} verification with proof assistants. In this
  context, the steps to verify that a transformation is semantic
  preserving include: (1) the formalization of the execution semantics
  of the source representation, (2) of the target representation, (3)
  the formal description of the transformation, and (4) the proof of a
  corresponding semantic preservation theorem. In this thesis, these
  steps have been carried within the framework of the \coq{} proof
  assistant.  Even though these steps are clearly set, the
  specificities of the \hilecop{} model-to-text transformation,
  compared to compilers for programming languages, bring some
  interesting research challenges.  Finally, we have brought the
  informal \textit{paper} proof that the \hilecop{} transformation is
  semantic preserving by demonstrating a related behavior preservation
  theorem.  The full mechanization of the proof using the \coq{} proof
  assistant is an ongoing task.
\end{abstract}

%----------------------------------------------------------------------------------------
%	FRENCH EXTENDED ABSTRACT
%----------------------------------------------------------------------------------------

\include{Chapters/FrenchExtendedAbstract}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

% \include{Chapters/Introduction}
% \include{Chapters/PrelimNotions}
% \include{Chapters/SITPN}
% \include{Chapters/H-VHDL}
% \include{Chapters/Transformation}
% \include{Chapters/Proof}
% \include{Chapters/Conclusion} 

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

% \include{Appendices/PlaceDesign}
% \include{Appendices/TransitionDesign}
% \include{Appendices/tree-beh-pres-thm}
% \include{Appendices/SemanticPreservationProof}


%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\printbibliography[heading=bibintoc]

%----------------------------------------------------------------------------------------

\end{document}  

